---
title: "Week 2 - Excercises"
author: "Annika Hirsch"
date: '2024-04-26'
output: 
  html_document:
    theme: flatly
    highlight: tango
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remark**: I prefer working with Markdown documents, since I think that they are easier to organise.

```{r Packages}
library("readr")
library("sf")
library("dplyr")
library("tmap")
library("tidyr")
library("ggplot2")
library("FITfileR")

#update.packages()
#install.packages("ggplot2")
```

# Excercise A

## Task 1: Data Import

```{r A Task 1}
wildschwein_BE <- read_delim("wildschwein_BE_2056.csv", ",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056)

head(wildschwein_BE)
```

## Task 2: Overview

```{r A Task2 TimeDiff}
difftime_secs <- function(later, now){
    as.numeric(difftime(later, now, units = "secs"))
}

now <- wildschwein_BE$DatetimeUTC[4]
later <- wildschwein_BE$DatetimeUTC[400]

print(difftime(later, now))
```
Now inspect your data in more detail. Try to answer the following questions:

* How many individuals were tracked?
* For how long were the individual tracked? Are there gaps?
* Were all individuals tracked concurrently or sequentially?
* What is the temporal sampling interval between the locations?

```{r A Task2 Questions, results = 'asis'}
# Number of individuals:
no_individuals <- length(unique(wildschwein_BE$TierID))
print(paste0("There were ", no_individuals, " individuals tracked."))

# Duration of tracking of the individuals

# store IDs of individuals:
indIDs <- unique(wildschwein_BE$TierID)

# Create a loop that calculates the individual time difference of the first and last entry of each individual
# I should probably use group by... but I've already done the loop.. 
for (i in indIDs){
  subset <- subset(wildschwein_BE, TierID == i)
  mintime <- min(subset$DatetimeUTC)
  maxtime <- max(subset$DatetimeUTC)
  timediff <- difftime(maxtime, mintime, units = "days")
  print(paste0("The individual with the id ", i, " was tracked over a time interval of ", timediff, " days."))
  # Add Functions to determine if there were gaps
  subset$timelag <- difftime_secs(lead(subset$DatetimeUTC), subset$DatetimeUTC)
  lag_days <- difftime(lead(subset$DatetimeUTC), subset$DatetimeUTC, units = "hours")
  maxlag <- max(na.omit(lag_days))
  #print(table(subset$timelag)) # Nice to have, but takes up a lot of space and information not really needed. 
  print(paste0("The maximum time lag (=gap) of individual ", i, " was ", maxlag, " hours."))
  plot(table(subset$timelag))
  # Display the start and end date of each individual to determine if they were recorded concurrently or sequential.
  print(paste0("The individual ", i, " was recorded from ", mintime, " to ", maxtime))
}

# Add timelag column
wildschwein_BE <- wildschwein_BE %>%
  group_by(TierID) %>%
  mutate(timelag = difftime_secs(DatetimeUTC, lag(DatetimeUTC)))

# Summary
wildschwein_BE %>%
  st_drop_geometry() %>%
  summarise(mean = mean(timelag, na.rm = TRUE), median = median(timelag, na.rm = TRUE))
```

There were three individuals tracked. They were tracked for 234.67, 262.66 and 338.58 days. The biggest time gap was 16.77 hours. Considering the timeframe of more than 230 days, these 16 hours do not seem much. The individuals were tracked concurrently. The median of the time lag is around 900 seconds (15 mins). Therefore the temporal sampling interval is roughly 15 minutes. 

## Task 3: Distance between locations

```{r A Task3 Distance}
later <- lag(wildschwein_BE$geometry)
now <- wildschwein_BE$geometry

#st_distance(later, now, by_element = TRUE)

distance_by_element <- function(later, now){
  as.numeric(
    st_distance(later, now, by_element = TRUE)
  )
}

wildschwein_BE <- wildschwein_BE %>% 
  group_by(TierID) %>%
  mutate(
    steplength = distance_by_element(lag(geometry), geometry)
        )

wildschwein_BE %>%
  st_drop_geometry() %>%
  summarise(mean = mean(steplength, na.rm = TRUE), median = median(steplength, na.rm = TRUE))
  
```

## Task 4: Deriving distance & speed

```{r A Task4 speed}
wildschwein_BE$speed <- wildschwein_BE$steplength / wildschwein_BE$timelag
```

## Task 5: Plausibility Check

```{r A Task5 Plausibility}
wildschwein_sample <- wildschwein_BE %>%
  filter(TierName == "Sabi") %>%
  head(100)

tmap_mode("view")

tm_shape(wildschwein_sample) + 
  tm_dots()
```

```{r A Task5 plot}
wildschwein_sample_line <- wildschwein_sample %>%
  # dissolve to a MULTIPOINT:
  summarise(do_union = FALSE) %>% 
  st_cast("LINESTRING")

tmap_options(basemaps = "OpenStreetMap")

tm_shape(wildschwein_sample_line) +
  tm_lines() +
  tm_shape(wildschwein_sample) + 
  tm_dots()
```

# Excercise B

```{r B data}
caro <- read_delim("caro60.csv", ",") %>%
  st_as_sf(coords = c("E","N"), crs = 2056) %>% 
  select(DatetimeUTC)
```

## Task 1: Calculate speed at scale 1

```{r B Task1}
# Timelag
caro <- caro %>%
  mutate(timelag1 = difftime_secs(lead(DatetimeUTC), lag(DatetimeUTC)))

# Steplength
caro <- caro %>% 
  mutate(steplength1 = distance_by_element(lag(geometry), lead(geometry)))

# Speed
caro$speed1 <- caro$steplength1 / caro$timelag1

# View: 
head(caro)
```

## Task 2: Calculate speed at scale 2

```{r B Task2}
# Timelag
caro <- caro %>%
  mutate(timelag2 = difftime_secs(lead(DatetimeUTC, n = 2), lag(DatetimeUTC, n = 2)))

# Steplength
caro <- caro %>% 
  mutate(steplength2 = distance_by_element(lag(geometry, n = 2), lead(geometry, n = 2)))

# Speed
caro$speed2 <- caro$steplength2 / caro$timelag2

# View: 
caro %>%
  st_drop_geometry() %>% 
  select(timelag2, steplength2, speed2) %>% 
  head()
```

## Task 3: Calculate speed at scale 3

```{r B Task3}
# Timelag
caro <- caro %>%
  mutate(timelag3 = difftime_secs(lead(DatetimeUTC, n = 4), lag(DatetimeUTC, n = 4)))

# Steplength
caro <- caro %>% 
  mutate(steplength3 = distance_by_element(lag(geometry, n = 4), lead(geometry, n = 4)))

# Speed
caro$speed3 <- caro$steplength3 / caro$timelag3

# View: 
caro %>%
  st_drop_geometry() %>% 
  select(timelag3, steplength3, speed3) %>% 
  head()

```

## Task 4: Compare speed across scales

```{r B Task4 long table}
caro2 <- caro %>% 
  st_drop_geometry() %>% 
  select(DatetimeUTC, speed1, speed2, speed3)

caro_long <- caro2 %>%
  pivot_longer(c(speed1, speed2, speed3))
  
head(caro_long)
```

```{r B Task4 plot}
ggplot(caro_long, aes(name, value)) +
  geom_boxplot(outliers = FALSE)
```

# Excercise C

```{r C loading required package, eval = FALSE}
#if(!requireNamespace("remotes")) {
#    install.packages("remotes")
#}
#remotes::install_github("grimbough/FITfileR")
```

```{r C activity data import}
activity <- readFitFile("kurz_auf_den_harder.fit")

activity

activity_records <- records(activity)
activity_records

# Does not work the way I want it to. Does not add the distance to the correct rows. 
#activity_records <- records(activity) %>% 
#  bind_rows() %>% 
#  arrange(timestamp)

#activity_records
```

```{r C mapping}
# Transform into sf
activity_sf <- st_as_sf(activity_records[["record_1"]],
                        coords = c("position_long", "position_lat"),
                        crs = 4326
)

# Change crs
st_transform(activity_sf, crs = 2056)

tmap_mode("view")

# Create a line
activity_line <- activity_sf %>%
  summarise(do_union = FALSE) %>% 
  st_cast("LINESTRING")

# Change basemap
tmap_options(basemaps = "OpenStreetMap")

# Plot map
tm_shape(activity_line) +
  tm_lines() #+ # Points are really dense...
  #tm_shape(activity_sf) + 
  #tm_dots()
```


